More on the Slavnov products of spin chains, KP hierarchy tau and Baker-Akhiezer functions

### 2025-05-13 draft.tex
improve section 2 of the draft

There are two important consequences for us now. First of all, if we
assume that this system has a non-trivial solution, it will be written
as minors of the matrix \(\bm{M}\).

Moreover, since the system is homogeneous, the solutions are
determined up to multiplicative terms, that we can fix them by
requiring that the final result gives the required tau function. In
the original works of Foda and collaborators, e.g.~\cite{Foda:2009zz,
  Wheeler:2010vmq}, the authors perform a series of redefinitions to
reach the same goal.

Evidently being a determinant is not enough to guarantee that the
Slavnov product is a tau function.  In order to accomplish this
result, we need to specify the models. Following the original
work~\cite{Belliard:2019bfz}, we will only discuss the rational models
-- we will see that the calculations are overwhelming even in this
simple case. The generalization for open boundary conditions and for
trigonometric models is immediate, see~\cite{Slavnov:2019aba} for more
details on the generalization.

\subsection{Rational models}

Let us startwith an R-matrix of the form
\begin{equation}
  R(u,v) = \mathbb{1} + g(u,v) \mathcal{P}\; , \qquad
  g(u,v) = \frac{c}{u - v}
\end{equation}
where \(c\) is a constant, the identity operator is denoted by
\(\mathbb{1}\) and \(\mathcal{P}\) is the permutation operator.

In the models with rational R-matrix, the eigenvalues of the transfer matrix
have the form 
\begin{equation}
  \Lambda(z, \bm{v}) = g(z, \bm{v}) \mathcal{Y}(z; \bm{v})\; ,
\end{equation}
where the function \(\mathcal{Y}(z; \bm{v})\) is symmetric over the
Bethe roots \(\bm{v}\), and linearly depends on every
\(v_j\).

Generically, it is written as
\begin{equation}
  \mathcal{Y}(z; \bm{v}) = \sum_{p=0}^n \alpha_p(z) \sigma_p^{(n)}(\bm{v})\; , 
\end{equation}
where \(\sigma_p^{(n)}(\bm{v})\) are elementary symmetric polynomials
in \(\bm{v}\) and where \(\alpha_p(z)\) are free functional
parameters. In fact, in the XXX discussed in~\cite{Belliard:2019bfz},
these functions all \(\alpha_p(z)\) are polynomials of degree \(2n\).
The important point for us is that the functions \(\alpha_p(z)\) are
not singular when we consider the limit \(z \to v_j \in \bm{v}\), and
that also means that \(\mathcal{Y}(z; \bm{v})\) is also regular in
this limit.

Let us also define the following product 
\begin{equation}
  g(z, \bm{v}) = \prod_{v_i\in \bm{v}} g(z; v_i)\; , 
\end{equation}
and that means that each Bethe root \( v_i \in \bm{v}\) 
is a pole of \(g(z;\bm{v})\). Therefore, if \(\Gamma\) is a contour
containing all the Bethe roots \(\{v_j\}_{j=1}^n\), we have 
\begin{equation}
  \frac{1}{2 \pi i}\oint_{\Gamma} \mathrm{d}z g(z; \bm{v}) = \sum_{j=1}^n  \prod_{v_i \in \bm{v}\setminus \{u_j\}} g(z, v_i)\; .
\end{equation}
Furthermore, the coefficients \(L_{jk}\) are given by
\begin{equation}
  L_{jk} = g(u_k, \bm{u}_k) \mathcal{Y}(u_k; \bm{u}_j)
\end{equation}
and from this expression one can calculate the matrix \(\bm{M}\).

It has also been shown~\cite{Belliard:2019bfz} that if \(\det
(\bm{M}) = 0 \) and \(\mathrm{rank}(\bm{M}) = n\), then, the Slavnov
products are given by
\begin{equation}
 \bm{\zeta}_\ell = \phi(\bm{v}) \tilde{\Delta}(\bm{u}_\ell)\hat{\Omega}_\ell\; ,
\end{equation}
where \(\phi(\bm{v})\) is a function of the Bethe roots and
\begin{equation}
  \tilde{\Delta}(\bm{u}_\ell) = \prod_{\substack{u_j, u_j \in \bm{u}_\ell \\ j>k}} \frac{c}{u_j - u_k}\quad \Rightarrow \quad
\frac{\tilde{\Delta}(\bm{u}_\ell)}{c_0} = \prod_{\substack{u_j, u_j \in \bm{u}_\ell \\ j>k}} \frac{1}{u_j - u_k}\; , 
\end{equation}
where we have absorbed the product of \(c\) into a new constant \(c_0\). 
All in all, \(\tilde{\Delta}(\bm{u}_\ell)/c_0\) is the inverse of the Vandermonde determinant
\(\Delta(\bm{u}_\ell) = \prod_{j>k}(u_j - u_k)\).

>>>>

There are two important consequences for us now. First of all, if we
assume that this system has a nontrivial solution, it will be
expressed in terms of minors of the matrix \(\bm{M}\).

Moreover, since the system is homogeneous, the solutions are
determined up to multiplicative factors, which we can fix by
requiring that the final result yields a KP tau function. In the
original works of Foda and collaborators, e.g.,~\cite{Foda:2009zz,
  Wheeler:2010vmq}, the authors perform a series of redefinitions to
achieve the same goal.

Evidently, being a determinant is not enough to guarantee that the
Slavnov product is a tau function.  In order to establish this
result, we need to specify the models. Following the original
work~\cite{Belliard:2019bfz}, we will restrict the current analysis to rational models
-- we will see that the calculations are overwhelming even in this
simple case. The generalization to open boundary conditions and for
trigonometric models is straightforward, see~\cite{Slavnov:2019aba} for more
details.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Rational models}

Let us start with an R-matrix of the form
\begin{equation}
  R(u,v) = \mathbb{1} + g(u,v) \mathcal{P}\; , \qquad
  g(u,v) = \frac{c}{u - v}
\end{equation}
where \(c\) is a constant, \(\mathbb{1}\) is the identity and
\(\mathcal{P}\) is the permutation operator~\cite{Slavnov:2019aba}.
The eigenvalues of the transfer matrix have the form
\begin{equation}
  \Lambda(z, \bm{v}) = g(z, \bm{v}) \mathcal{Y}(z; \bm{v})\; ,
\end{equation}
where the function \(\mathcal{Y}(z; \bm{v})\) is symmetric over the
Bethe roots \(\bm{v}\), and has a linear dependence on each Bethe root
\(v_j\).

Generically, we write the \(\mathcal{Y}\)-functions as 
\begin{equation}
  \mathcal{Y}(z; \bm{v}) = \sum_{p=0}^n \alpha_p(z) \sigma_p^{(n)}(\bm{v})\; , 
\end{equation}
where \(\sigma_p^{(n)}(\bm{v})\) are elementary symmetric polynomials
in \(\bm{v}\), and \(\alpha_p(z)\) are free functional parameters.
The most important point for us is that the functions \(\alpha_p(z)\), 
and consequently the \(\mathcal{Y}\)-function, are regular in the limit
\(z \to v_j \in \bm{v}\). The XXX spin chain has been discussed
in~\cite{Belliard:2019bfz}, and the authors show that \(\alpha_p(z)\)
are polynomials of degree \(2n\).

Let us also define the product 
\begin{equation}
  g(z, \bm{v}) = \prod_{v_i\in \bm{v}} g(z; v_i)\; . 
\end{equation}
Now, it is easy to see that each Bethe root \( v_i \in \bm{v}\) is a
pole of \(g(z;\bm{v})\). Therefore, if \(\Gamma\) is a contour
containing all the Bethe roots \(\{v_j\}_{j=1}^n\), we have
\begin{equation}
  \frac{1}{2 \pi i}\oint_{\Gamma} \mathrm{d}z g(z; \bm{v}) = \sum_{j=1}^n
  \prod_{v_i \in \bm{v}\setminus \{v_j\}} g(z, v_i)\; .
\end{equation}
Putting all these facts together, the coefficients \(L_{jk}\) are given by
\begin{equation}
  L_{jk} = g(u_k, \bm{u}_k) \mathcal{Y}(u_k; \bm{u}_j)\; ,
\end{equation}
and from this expression one can calculate the matrix \(\bm{M}\).

It has also been shown~\cite{Belliard:2019bfz} that if \(\det (\bm{M})
= 0 \) and \(\mathrm{rank}(\bm{M}) = n\), the Slavnov products are
given by
\begin{equation}
 \bm{\zeta}_\ell = \phi(\bm{v}) \tilde{\Delta}(\bm{u}_\ell)\hat{\Omega}_\ell\; . 
\end{equation}
Let us now explain the different terms in this expression. 
First of all, \(\phi(\bm{v})\) is a function of the Bethe roots and
\begin{equation}
  \tilde{\Delta}(\bm{u}_\ell) = \prod_{\substack{u_j, u_j \in \bm{u}_\ell \\ j>k}} \frac{c}{u_j - u_k}\; .
\end{equation}
We can absorbed the product of \(c\) into a new constant
\(c_0\), and it is easy to see that \(\tilde{\Delta}(\bm{u}_\ell)/c_0\) is the
inverse of the Vandermonde determinant \(\Delta(\bm{u}_\ell) =
\prod_{j>k}(u_j - u_k)\).
